---
title: "Homework 24"
format: html
editor: visual
---

## Overview

For your homework you will implement a data analysis using a dataset on crab growth measurements.

Set things up (hint: there is a new package to add with `library()`)

```{r}
rm(list = ls())
library(tidyverse)
library(here)
library(ggfortify)
```

## The data

Let's pull in some data appropriate biologists: crab measurements. This dataset includes measurements of farmed crabs.

```{r}
crabs <- read.csv(here("data", "CrabAgePrediction (1).csv"))
glimpse(crabs)
```

One of the variables is categorical - let's fix that in the data.

```{r}
crabs$Sex <- as.factor(crabs$Sex)
```

## Question

Is there a relationship between height and weight in female crabs?

1.  Extract the data for just female crabs.

    ```{r}
    f_crabs <- crabs |>
      filter(Sex == "F")

    glimpse(f_crabs)
    ```

2.  Examine the response and predictor variables and use the flow chart from class to decide the appropriate statistical analysis (choose from: chi-square, 2-sample t-test or linear regression). Explain your answer here:

    The response and predictor variables are both continuous ratio data so you should use a linear regression model.

3.  Work through the data analysis **following all of the steps provided in chapter 5 of your book**. You may use class notes or homework keys as a guide.

    ### Plot the data

    ```{r}
    ggplot(f_crabs, aes(x = Height, y = Weight))+
      geom_point()
        
    ```

    There is one outlier that makes it very hard to estimate the slope or intercept

    ### Fit a model and check assumptions

    ```{r}
    f_crabs_model <- lm(Weight ~ Height, data = f_crabs)

    autoplot(f_crabs_model, smooth.color = NA)
    ```

The autoplot plots show a lot of problems with the using a linear model to represent this data.

The residuals vs fitted plot does not follow the dashed line, so a line may not be appropriate to fit the model.

The Normal Q-Q plot almost follows the dotted line signifying a normal distribution but one outlier is very far from the dotted line.

The scale-location plot is also pretty weird and means there may not be constant variance.

The residuals vs leverage plot shows that one point has a lot of leverage on the data.

### Interpretation of the linear model

```{r}
anova(f_crabs_model)
summary(f_crabs_model)
```

There seems to be a significant relationship between the height and weight of female crabs (F = 613.57, DF = 1, P \< 0.05).

### add a regression line to the plot

```{r}
ggplot(f_crabs, aes(x = Height, y = Weight))+
  geom_point()+
  geom_smooth(method = 'lm')
```

regression line enforces the positive relationship between height and weight in female crabs.
