---
title: "Biostats - Homework 20"
format: html
editor: visual
---

### Introduction

The objective of this assignment is to have you continue practicing the skills you have learned in Chapters 3 and 4 of Getting Started with R and reinforce the ideas we discussed in class regarding the Weissgerber et al. paper. Please complete all of the following steps:

1.  Write the code to set up your workspace.

```{r}
rm(list = ls())
library(tidyverse)
library(here)
```

2.  Read in the dataset called soil_data.csv and name it "soil."

```{r}
soil <- read.csv(here("data", "soil_data.csv")) 
soil |> View()
```

3.  Create a histogram for the SOM_percent column in the dataframe. Accept the default number of bins.

```{r}
ggplot(soil, aes(x = SOM_percent))+
  geom_histogram()
```

How would you describe the mean_SOM data based on viewing the histogram? Write your answer below:

The mean_SOM seems like it is relatively low becuase the data is heavily skewed to the right.

4.  Create a new dataframe called SOM by calculating for each forest the mean soil organic matter and the standard deviation of soil organic matter. Make sure to name variables you create "mean_SOM" and "sd_SOM" so that the next code chunk I give you works.

```{r}
SOM <- soil |>
  group_by(Forest) |>
  summarise(mean_SOM = mean(SOM_percent),
            sd_SOM = sd(SOM_percent))
```

5.  Now run the following code chunk which will create a bar plot of mean soil organic matter per forest, with standard deviation as error bars.

```{r}
ggplot(SOM, aes(Forest, mean_SOM))+
  geom_errorbar(aes(ymax = mean_SOM + sd_SOM, ymin = mean_SOM - sd_SOM), width = 0.2)+
  geom_col(fill = "white", color = "black")+
  ylab("Mean percent soil organic matter")+
  xlab("Forest name")
```

Based on what you see in the bar chart, how would you describe the pattern of SOM between the 4 different forests? Write your answer here:

The SOM seems to be higher on average i the Kip Tract and Donnerville forests, however, those forests also seem to have more variability in SOM.

Now create a boxplot of the SOM_percent data. Use Forest as the x-axis variable and overlay the point data on top of the boxplot. Make the points blue and somewhat transparent. Call the x-axis "Forest name" and the y-axis "Percent soil organic matter" (hint: See homework 18...).

```{r}

ggplot(soil, aes(x = Forest, y = SOM_percent))+
  geom_boxplot()+
  geom_point(color = "blue", alpha = 0.5)+
  labs(x = "Forest Name",
       y = "Percent Soil Organic Matter")
```

Based on what you see in the box plot, write out your answers the following questions:

-   how would you describe the pattern of SOM between the 4 different forests?

    The average SOM is relatively low in the Degrasse, Donnerville, and South Hammond forests, while it is higher in the Kip Tract. The SOM is much more variable in the Kip Tract than in the other three forests, however there does seem to be a high concentration of points in the low end of the boxplot.

-   What do you notice that is different from how the data were displayed in the bar plot?

    You can see the distribution of the points much more easily. You can also see that the Donnerville forest is similar in SOM to the South Hammond and Degrasse forests, however it has one outlier that made it appear higher on the bar plot.

-   Which plot type is more informative? Explain your answer.

    The boxplot is much more informative because you can see more of the data, rather than just the mean value and the standard deviation for each site. The boxplot better accounts for outliers and variation in the dataset than the bar plot.
