---
title: "Homework-21"
format: html
---

### Set Up

```{r}
rm(list = ls())
library(tidyverse)
library(here)
```

### Read in Datasets

```{r}
lady <- read.csv(here("data","ladybirds.csv"))
glimpse(lady)

ozone <- read.csv(here("data", "ozone.csv"))
glimpse(ozone)
```

### Make variables factors

```{r}
lady$Habitat <- as.factor(lady$Habitat)
levels(lady$Habitat)

lady$Site <- as.factor(lady$Site)
levels(lady$Site)

lady$colour <- as.factor(lady$colour)
levels(lady$colour)

ozone$Garden.location <- as.factor(ozone$Garden.location)
levels(ozone$Garden.location)

ozone$Garden.ID <- as.factor(ozone$Garden.ID)
levels(ozone$Garden.ID)
```

### 5.2 X^2 contingency table analysis

5.2.2 Organizing data for plotting and analysis
```{r}
totals <- lady %>%
group_by(Habitat, colour) %>%
summarise(total.number = sum(number)) 

ggplot(totals, aes(x = Habitat, 
                   y = total.number, 
                   fill = colour)) +
geom_bar(stat = 'identity', position = 'dodge')

```

colors in the default chart aren't great, change manually:
```{r}
ggplot(totals, aes(x = Habitat, y = total.number, fill = colour)) +
geom_bar(stat = 'identity', position = 'dodge') + scale_fill_manual(values = c(black = "black", red = "red"))
```

making the X^2 test
```{r}
#first make a matrix
lady.mat <- xtabs(number ~ Habitat + colour, data = lady)
lady.mat

#run matrix through the test
chisq.test(lady.mat)

lady.chi <- chisq.test(lady.mat) 
names(lady.chi)

```

### 5.3 Two-sample t-test
a comparison of the means of two groups of numeric values.

plotting data
```{r}
ggplot(ozone, aes(x = Ozone)) + 
  geom_histogram(binwidth = 10) + 
  facet_wrap(~ Garden.location, ncol = 1) + theme_bw()
```
there may be a difference in the means of the two groups based off of the two graphs

```{r}
# Do a t.test now....
t.test(Ozone ~ Garden.location, data = ozone)
#returns a Welch t test, doesn't test for equality of variance

#t-test with test for equality of variance
var.test(Ozone ~ Garden.location, data = ozone)

```
in words: do ozone levels vary as a result of garden location

p value is low, indicates difference in means
confidence interval also doesn't include 0, indicating that it is very unlikely that the differenc in means in 0
