---
title: "homework-16"
format: html
---

#Set up

```{r}
rm(list = ls())
library(tidyverse)
library(here)
```

Read in the donnerville dataset

```{r}
cover <- read.csv(here("data", "donnerville_percent_cover.csv"))

#here I view the dataset 2 ways
glimpse(cover)
head(cover)
```

#changing variable types Site, plot type and cover type should be factors

```{r}
#making the Site variable into a factor
cover$Site <- as.factor(cover$Site)
levels(cover$Site) #levels are inconsistent and must be fixed

fix <- which(cover$Site %in% c("DONNERVILLE - 1", 
                               "donnerville 1",
                               "Donnerville 1"))


cover$Site[fix] <- "Donnerville 1"
cover$Site <- factor(cover$Site)

fix2 <- which(cover$Site %in% c("donnerville 2",
                                "Donnerville 2"))

cover$Site[fix2] <- "Donnerville 2"
cover$Site <- factor(cover$Site)

fix3 <- which(cover$Site %in% c("Donnerville 3",
                                "donnerville-3"))

cover$Site[fix3] <- "Donnerville 3"
cover$Site <- factor(cover$Site)

levels(cover$Site) #now there are 3 levels of Site
```

repeating the process for plot type and cover type

```{r}
cover$Plot.type <- as.factor(cover$Plot.type)
levels(cover$Plot.type) #Plot.type already has two levels so it doesn't need fixed

cover$cover_type <- as.factor(cover$cover_type)
levels(cover$cover_type) #does need fixed

fix4 <- which(cover$cover_type %in% c("bare",
                                      "Bare"))

cover$cover_type[fix4] <- "Bare"
cover$cover_type <- factor(cover$cover_type)

fix5 <- which(cover$cover_type %in% c("bryophtes",
                                      "Bryophytes"))

cover$cover_type[fix5] <- "Bryophytes"
cover$cover_type <- factor(cover$cover_type)

fix6 <- which(cover$cover_type %in% c("vasc ",
                                      "Vascular"))

cover$cover_type[fix6] <- "Vascular"
cover$cover_type <- factor(cover$cover_type)

levels(cover$cover_type)
```

#make date into a date variable
```{r}
cover$Date <- dmy(cover$Date)
```

#Determine the average percent cover by cover type (bare, bryophytes, vascular plants) for Donnerville 1.
```{r}
cover |> 
  filter(Site == "Donnerville 1") |>
  group_by(cover_type) |>
  summarise(mean(Percent))
```

#Determine the average percent vascular plant cover across all sites.
```{r}
cover |>
  filter(cover_type == "Vascular") |>
  summarise(mean(Percent))
```

