---
title: "homework-18"
format: html
---

# GSR Chapter 4

set up quarto document

```{r}
rm(list = ls())
library(tidyverse)
library(here)
```

read in the compensation dataset

```{r}
compensation <- read.csv(here("data", "compensation.csv"))
```

## Using ggplot2

give dataframe to the ggplot function, variables to the aesthetic function (`aes()`)

Added arguments and features to the plot as they were presented in the textbook reading
```{r}
ggplot(compensation, aes(x = Root, y = Fruit, color = Grazing)) + 
  geom_point(size = 5) +
  xlab("Root Biomass") +
  ylab("Fruit Production") +
  theme_bw() #removes grey background

```

can also distinguish levels with a shape argument
```{r}
ggplot(compensation, aes(x = Root, y = Fruit, shape = Grazing)) + 
  geom_point(size = 5) +
  xlab("Root Biomass") +
  ylab("Fruit Production") +
  theme_bw() #removes grey background
```

## box and whisker plots

use `geom_boxplot()`
```{r}
ggplot(compensation, aes(x = Grazing, y = Fruit)) + geom_boxplot() +
xlab("Grazing treatment") +
ylab("Fruit Production") +
theme_bw()

```

can add raw data to the boxplot with `geom_point`
also adds a few more customization options (alpha = transparency)
```{r}
ggplot(compensation, aes(x = Grazing, y = Fruit)) + geom_boxplot() +
geom_point(size = 4, colour = 'lightgrey', alpha = 0.5) + xlab("Grazing treatment") +
ylab("Fruit Production") + theme_bw()

```

Do plants with wider root diameters at the start of the experiment produce more fruit?
  Plants with wider root diameters appear to produce more fruit.

Do grazed or ungrazed plants produce more fruit?
  Grazed plants produce more fruit.
  
## making histograms

use `geom_histogram()` to make a histogram in ggplot2. 
use the bins and bins_width arguements to adjust the width of bars
```{r}
ggplot(compensation, aes(x = Fruit)) + 
  geom_histogram(bins = 10)

```
use the bins and bins_width arguements to adjust the width of bars

## facetting
ggplot2 uses facetting to create lattice graphs
```{r}
ggplot(compensation, aes(x = Fruit)) + geom_histogram(binwidth = 15) + facet_wrap(~Grazing)
```

use `ggsave()` to download plots
```{r}
ggsave("ThatCoolHistogramOfFruit.png")
```

