---
title: "homework-23"
format: html
---

### Set Up

```{r}
rm(list = ls())
library(tidyverse)
library(here)
```

### 5.4 Intro linear models

-   General linear models are a class of model that include regression, multiple regression, ANOVA, and ANCOVA.

-   Use least squares framework for estimation and share a common set of assumptions

-   `lm()` function fits linear models

### 5.5 Simple linear regression

read in the data

```{r}
plant_gr <- read.csv(here("data", "plant.growth.rate.csv"))
glimpse(plant_gr)
```

plot the data

```{r}
ggplot(plant_gr, aes(x = soil.moisture.content,
                     y = plant.growth.rate))+
  geom_point()+
  ylab("Plant Growth Rate (mm/week)")+
  theme_bw()
  
```

By observing the plot you can estimate the slope and intercept before doing statistical analysis.

-   slope \~ 15

-   intercept is somewhere bewteen 15 and 20 mm/week

Now use `lm()` to fit a model

```{r}
model_pgr <- lm(plant.growth.rate ~ soil.moisture.content,
                data = plant_gr)
```

To check the assumptions of a linear model, use the `autoplot()` function from the ggfortify package.

```{r}
#install.packages("ggfortify")
library(ggfortify)

autoplot(model_pgr, smooth.color = NA)
```

`autoplot()` plots

1.  Residuals vs Fitted - tells us if a line is appropriate to fit the model, blue line should be as close to the dotted line as possible.

2.  Normal Q-Q - evaluates assumption of normality residual plot should follow the dotted line.

3.  Scale-Location - evaluates assumption of constant variance.

4.  Residuals vs Leverage - evaluates leverage, detects data points that may influence the dataset more than others

Interpretation of the linear model

```{r}
#anova function returns a sum of squares table, it does not perform an ANOVA that compares means
anova(model_pgr)

summary(model_pgr)
```

Large F-value means that error variance is small relative to variance caused by the explanatory variable.

Very small p-value means the pattern we observe isn't a chance result.

Intercept is 19.348

Slope is 12.750

### Add a regression line to a plot

```{r}
ggplot(plant_gr, aes(x = soil.moisture.content,
                     y = plant.growth.rate))+
  geom_point()+
  geom_smooth(method = 'lm')+ #adds regression line
  ylab("Plant Growth Rate (mm/week)")+
  theme_bw()
```
